### 2021-05-24 assertions/doc/thesis-bernhardt/chapters/5-offline_design_tezos.tex
Final revision - evaluation chpt 5

\section{Cost Analysis and Evaluation}\label{sec:cost_analysis_distributed}
In this section, the costs of checking a formula with the proposed approach and implementation for distributed assertion checking are evaluated. \secref{sec:usecase_cost} already determined the cost for a local validation as a benchmark. It stated \eqref{eq:gas_transaction} to calculate the gas consumption of a transaction, including internal operations. It identified the gas consumption of deserialization, data parsing and interpretation as growing factors, as they depend on the size of the input parameter.

According to the orchestration scheme, each validator executes $t+2$ operations with the same input parameter $p$. Since they apply lazily, the deserialization cost for the parameter and assertion code only have to be paid once. In addition to the interpretation cost, the reading, parsing and base costs have to be paid for each internal operation. 
\begin{align}\label{eq:gas_validator}
\begin{split}
GasConsumed_{validator}(p) > \quad t * (&base\, gas \\
&+ gasParsing(p) \\
&+ gasRead(C_{assertion}) \\
&+ gasParsing(C_{assertion}) \\
&+ gasInterpretation(C_{assertion})) \\
\end{split}
\end{align}
Based on the resulting inequality in \eqref{eq:gas_validator}, which specifies the costs incurred by each internal operation to the assertion contract as a basis of the total gas consumption per validator, it can be concluded that the overall gas consumption of the distributed validation is higher than in the case of a local validation.

The distributed approach was benchmarked by implementing the orchestration scheme for the two minimal dummy contracts \texttt{Foo} and \texttt{Bar} given in \secref{sec:usecase_cost}. Appendix \ref{apx:foobar_distributed} contains the Michelson code for the controller, assertion and parent contracts. In order to preserve comparability between benchmark of the local and the distributed implementation, the controller contracts were configured with a probability threshold $c = \frac{1}{e}$, s.t. that the necessary number of test runs is equal to the sizes of the search spaces. \figref{fig:cost_distributed} shows the reported gas consumption per validator on the test network Delphinet for different input sizes. Since the effective number of test runs is always a multiple of 32, the cost function approximates a step function.

@@

The results show, that the cost incurred by a single validator, checking a subset of a search space with size $n$, is already considerably higher than a complete validation in a local approach (cf. \figref{sec:usecase_cost}). In the case of the list parameter (contract \texttt{Bar}), the gas consumption even increases exponentially due to the multiplied reading and parsing costs. Moreover, the probability of accepting an invalid parameter with the given configuration is, in the worst case, 37\%, which results in a bad cost-performance ratio.

In conclusion to these results, the proposed implementation of the distributed scheme has to be evaluated as inefficient and unable to achieve an improvement regarding the scalability of applications on Tezos. However, the approach is nonetheless promising, as it reduces the required computation per node to a fraction of the verification, and thus also reduces the computational footprint of the transaction within the network. As a prerequisite, it is crucial to lower the costs of the verification. This thesis already mentioned some alternative approaches and implementations, which might achieve this; \secref{sec:alt_random} introduces the possibility to coordinate the validators in a systematic iteration of the search space. Although this does not reduce the cost per test run, this reduces the quantify of necessary test runs to achieve a reliable result. This chapter described several other orchestration schemes, some of which reduce the number of emitted internal operations. As the inter-contract communication accounts for most of the gas consumption in the modular approach, avoiding them is an effective way of reducing the costs. In this regard, a concrete improvement of the proposed architecture could be to merge the controller and assertion code into a single contract. This corresponds to the inverse hybrid architecture as described in \secref{sec:monolithic}. Instead of $(t+1)$ internal operations, this only requires a single internal operation calling the parent contract.

>>>>

\section{Cost Analysis and Evaluation}\label{sec:cost_analysis_distributed}
In this section, the costs of checking a formula, with the proposed approach and implementation for distributed assertion checking, are evaluated. \secref{sec:usecase_cost} already determined the cost for a local validation as a benchmark. It stated \eqref{eq:gas_transaction} to calculate the gas consumption of a transaction including internal operations. It identified the gas consumption of deserialization, data parsing and interpretation as growing factors, as they depend on the size of the input parameter.

According to the orchestration scheme, each validator executes $t+2$ operations with the same input parameter $p$. Since they apply lazily, the deserialization cost for the parameter and assertion code only have to be paid once. In addition to the interpretation cost, the reading, parsing and base costs have to be paid for each internal operation. 
\begin{align}\label{eq:gas_validator}
\begin{split}
GasConsumed_{validator}(p) > \quad t * (&base\, gas \\
&+ gasParsing(p) \\
&+ gasRead(C_{assertion}) \\
&+ gasParsing(C_{assertion}) \\
&+ gasInterpretation(C_{assertion})) \\
\end{split}
\end{align}
Based on the resulting inequality in \eqref{eq:gas_validator}, which specifies the costs incurred by each internal operation invoking the assertion contract as a basis of the total gas consumption per validator, it must be concluded that the overall gas consumption of the distributed validation is higher than in the case of a local validation.

The distributed approach was benchmarked by implementing the orchestration scheme for the two minimal dummy contracts \texttt{Foo} and \texttt{Bar} given in \secref{sec:usecase_cost}. Appendix \ref{apx:foobar_distributed} contains the Michelson code for the controller, assertion and parent contracts. In order to preserve comparability between the benchmark of the local and the distributed implementation, the controller contracts were configured with a probability threshold $c = \frac{1}{e}$, s.t. that the necessary number of test runs is equal to the sizes of the search spaces. \figref{fig:cost_distributed} shows the reported gas consumption per validator for different input sizes on the test network Delphinet. Since the effective number of test runs is always a multiple of 32, the cost functions approximate a step function.

@@

The results show, that the cost incurred by a single validator, checking a subset of a search space of size $n$, is already considerably higher than a complete validation in a local approach (cf. \figref{sec:usecase_cost}). In the case of the list parameter (contract \texttt{Bar}), the gas consumption even increases exponentially due to the multiplied reading and parsing costs. Moreover, the cost-performance ratio is also worse, as the probability of accepting an invalid parameter with the given configuration is 37\% in the worst case.

In conclusion to these results, the proposed implementation of the distributed scheme has to be evaluated as inefficient and unable to achieve an improvement regarding the scalability of applications on Tezos. However, by reducing the required computation per node to a fraction of the verification, it has the potential to reduce the computational footprint of the transaction within the network. As a prerequisite, it is crucial to lower the costs of the verification. This thesis already mentioned some alternative approaches and implementations, which might achieve this; \secref{sec:alt_random} introduces the possibility to coordinate the validators in a systematic iteration of the search space. Although this does not reduce the cost per test run, it reduces the quantity of necessary test runs to achieve a reliable result. Furthermore, this chapter described several other orchestration schemes, some of which reduce the number of emitted internal operations. As the inter-contract communication accounts for most of the gas consumption in the modular approach, avoiding them is an effective way of reducing the costs. In this regard, a concrete improvement of the proposed architecture could be to merge the controller and assertion code into a single contract. This corresponds, to some extend, to the inverse hybrid architecture as described in \secref{sec:monolithic}. Instead of $t+1$ internal operations, this only requires a single internal operation calling the parent contract (provided that the controller an assertion code are merged into one entrypoint).
